---
title: "Introduction to golmDB"
author:
  - name: Thomas Wilson
    affiliation: Institute of Biological, Environmental & Rural Sciences (IBERS), Aberystwyth University, UK
    email: tpw2@aber.ac.uk
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction_to_golmDB}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The Golm Metabolome Database is a collection of reference mass spectra from biologically active metabolites that have been quantified using Gas Chromatography (GC) Mass Spectrometry (MS). The `golmDB` package provides a simple method to parse the Golm database, so it can be easily used in downstream processing pipelines. 

## Installation

`golmDB` can be installed directly from GitHub using the `remotes` package. 

```{r, install, echo=TRUE, eval=FALSE}
remotes::install_github('wilsontom/golmDB')
```

## Creating a golmDB object

Before creating a new `golmDB` object, you first need to download the version of Golm that you require. All the different Golm database files can be downloaded [here](http://gmd.mpimp-golm.mpg.de/download/). `golmDB` requires the database to be downloaded in the **MSP** format. 

Once you have the Golm database file, you can create a new `golmDB` object by using the `openGolmDB` function with the path of the database file. 

```{r, open, echo=TRUE, eval=TRUE}
library(golmDB)

GolmDB_MDN35 <-
  openGolmDB('../inst/extdata/GMD_20111121_MDN35_FAME_MSP.txt')

GolmDB_MDN35
```

`golmDB` utilises two separate S4 classes. The `GolmDB` class contains the entire database; where each database accession is stored in it's own `GolmAccession` class. The `GolmDB` accession slot, is large list of `GolmAccession` objects. 


```{r, slots, echo=TRUE, eval=TRUE}
slotNames(GolmDB_MDN35)

length(GolmDB_MDN35@accessions)

GolmDB_MDN35@accessions[[1]]

```


## Search Database

The `searchDB` method can be used to create smaller database objects, based on the results of searching for a particular pattern. For example, if you are only interested in Fatty Acid Methyl Esters then you can use 'methyl esters' as a search pattern. 

```{r searchDB, echo=TRUE, eval=TRUE}
GolmDB_FAME <- searchDB(GolmDB_MDN35, pattern = 'methyl ester')

GolmDB_FAME
```





